---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(lubridate)
library(tsibble)
library(tsibbledata)
library(fable)
library(feasts)

```



Load in the nyc_bikes data from the tsibbledata package. Have an initial look at it to see what you’re working with. Create three new columns: one that stores only the year the bike was used, one that stores only the month the bike was used, and one that stores the date. Use the data stored in start_time to create these new columns.

```{r}
nyc_bikes <- nyc_bikes

names(nyc_bikes)  
dim(nyc_bikes)
head(nyc_bikes,3)

```




```{r}
nyc_bikes_year <-
nyc_bikes %>%
  select(-gender, -type, -birth_year) %>%
  mutate(year = year  (start_time),
         month = month(start_time, label = TRUE),
         date  = date(start_time))


```


2 Question 2

Summarise the number of bike hire counts by month. Make a plot of this data. *Hint: remember that to group time series (tsibble) data, you need to use index_by instead of group_by before your summarise function. What does this plot tell you about the time series? Do you think this downsampled data would be adequate to build a forecast with?

```{r}

nyc_bikes_month <-
nyc_bikes_year %>%
index_by(month) %>%
  summarise(count = n())


  ggplot(nyc_bikes_month) +
  aes(x = month, y = count) +
  geom_point() + 
  geom_line(group = 1) +              
  ggtitle("Bike hire in NYC") + 
  xlab("Month") + ylab ("total count")
  

```

Question 3

Now Summarise the number of bike hire counts by date. Make a plot of this new aggregated data. What does this plot tell you about the time series? Would this data be preferrable for time series forecasting compared to the monthly data? 

```{r}

nyc_bikes_date <-
nyc_bikes_year %>%
index_by(date) %>%
  summarise(count = n())


    ggplot(nyc_bikes_date) +
  aes(x = date, y = count) +
  geom_point() + 
  geom_line(group = 1) +              
  ggtitle("Bike hire in NYC") + 
  xlab("Month") + ylab ("total count")




```

4 Question 4

Let’s begin to build a model. For this, we will use the downsampled by date dataset we created above in question 3. If you haven’t managed to get the code to do this, you can find it below.

```{r}
# fill the gaps with median

nyc_bikes_filled <- nyc_bikes_date %>%
  fill_gaps(count = as.integer(median(count)))

nyc_bikes_filled

```

```{r}

fit <- nyc_bikes_filled %>%
  model(
    snaive = SNAIVE(count),
    mean_model = MEAN(count),
    arima = ARIMA(count)
  )

fit

```

5 Question 5

Now we have our model fit, build a forecast to predict bike use over the next four months. Plot your models alongside your data.
Hint: forecast parameter would be roughly 120 (30 days x 4 months)

```{r}

forecast_bike_use <- fit %>%
  fabletools::forecast(h = 120)

forecast_bike_use



forecast_bike_use %>%
  autoplot(nyc_bikes_filled, level = NULL) +
  ggtitle("Forecasts for bikes usage") +
  xlab("month") +
  guides(colour = guide_legend(title = "Forecast"))
  

```



Question 6

Test your model accuracy : choose a training data set from your main dataset, build a forecast on the training set, and then plot the training set forecast against the real data. Calculate model accuracy.


```{r}

# Set training data from 2018-04-01 to 2018-08-31

train <- nyc_bikes_filled %>%
  filter_index("2018-04-01" ~ "2018-08-31") 


# run the model on the training set 
bikes_fit_test <- train %>%
  model(
    mean_model = MEAN(count),
    arima = ARIMA(count),
    snaive = SNAIVE(count))

bikes_fit_test

```


```{r}

# forecast from the training set
forecast_bike_use <- bikes_fit_test %>%
  fabletools::forecast(h = 12*6)

# Plot forecasts against actual values
forecast_bike_use %>%
  autoplot(train, level = NULL) +
    autolayer(nyc_bikes_filled, color = "black")

```


Question 7

Look at your forecast plots and accuracy values. Describe your results. Are your models a good fit for the data? If not, why not? What would you suggest doing with the data if you were expected to present these back to a client? For example, would you ask for more data? Would you test a different model?



A . SUMMARISE YOUR FINDINGS

Forecast isn’t great. This could be for several reasons:

    not enough data
    something is affecting it
    wrong model choice

I’d collect more data if possible - we only have a year’s worth so far.



























